---
title: "The effects of housing rehabilitation policies on residential mobility preferences in Andalusia"
output: html_document
date: "28-06-2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

This project is an examination of the RUCOSA survey made by The Urban Governance Lab in Andalusia to evaluate the effects of EU urban policies on vulnerable neighbourhoods

Its main hypothesis are the following:

$H_1$: The citizens in rehabilitated households have less propensity to leave neighbourhood (Individiual effect)

$H_2$: The citizens in the affected areas have less propensity to leave the neighbourhood, even if they have not been directly affected by these policies.

## 2. Data loading and cleaning

### 2.1. Packages and dataset
```{r, results='hide', warning=FALSE, message=FALSE, error=FALSE}
## necessary packages

library(tidyverse)
library(readr)
library(sem)
library(randomForest)
library(lmtest)
library(AER)
library(lavaan)
library(cowplot)
library(semPlot)
library(vtable)
library(ggdag)
library(dagitty)

# Save the dataset in "Downloads" directory in order to open it

USUARIO <- Sys.getenv("USERNAME")

df <- read_delim(paste0("C:/Users/",USUARIO,"/Downloads/RCS_PT3_270223_Indexes.csv"), delim = ";", escape_double = TRUE, trim_ws = TRUE)

```

### 2.2. Cleaning

```{r, results='hide', warning=FALSE}

# Years living in the neighbourhood

df$P0109 <- ifelse(df$P0109 == 98, df$C02, df$P0109)

df$P0109 <- ifelse(df$P0109 > 90, NA, df$0109)

# deal with NA

df[, !(names(df) %in% c("P0109"))] <- lapply(df[, !(names(df) %in% c("P0109"))], function(x) ifelse(x > 76, NA, x))

# Estimating NA for more than 30% missing is not considered:

df <- df[colSums(is.na(df))/nrow(df) < .30]

# character and numeric to factor

df <- df %>% mutate_if(is.character, as.factor)


df$TIPO_ZONA <- cut(df$TIPO_ZONA, breaks = c(0, 1, 2, 3), labels = c("Experimental", "Control", "Other"))

# Eliminate control variable:

df <- df[df$TIPO_ZONA != 'Other', ]

df <- df %>% 
  mutate(exp = ifelse(TIPO_ZONA == "Experimental", 1, 0))

# eliminate  variables with unique values

df <- select(df, -one_of(names(df)[sapply(df, function(x) n_distinct(x) == nrow(df))]))


sum(is.na(df$P0102_C))/length(df$P0102_C)

# median and mode AND other na impute
df <- na.roughfix(df)


df$P0209[df$P0209 > 1] <- 0
df$P0214[df$P0214 == 2] <- 0
df$P0219[df$P0219 == 2] <- 0


# building "social support" variable
df <- df %>% mutate(apoyo = log((P0905_A + P0905_B + P0905_C + P0905_D + P0905_E) / 5))

# for log(0)
df$apoyo[df$apoyo == -Inf] <- min(df$apoyo[is.finite(df$apoyo)])

# building "housing problems" variable:


df$P1009_A[df$P1009_A == 2] <- 0
df$P1009_B[df$P1009_B == 2] <- 0
df$P1009_C[df$P1009_C == 2] <- 0
df$P1009_G[df$P1009_G == 2] <- 0

df <- df %>% mutate(problemviv = (P1009_A + P1009_B + P1009_C + P1009_G) / 4)

# building "participation" variable (considered but not included)

var_names <- paste0("P0401_", LETTERS[1:10])

# Loop through the variable names and replace values
for (var_name in var_names) {
  df[[var_name]][df[[var_name]] < 4] <- 1
  df[[var_name]][df[[var_name]] == 4] <- 0
}

df <-  df %>% mutate(relvel= (P0401_A + P0401_B + P0401_C + P0401_D + P0401_E + P0401_F + P0401_G + P0401_H + P0401_I + P0401_J) / 10)

# building "increase in housing satisfaction variable"
df = df %>% mutate(diffviv =(P0224 - P0225))

# categories of families
df$P0902 <- cut(df$P0902,
                breaks = c(-Inf, 1, 2, 3, 4, 5, 6, 7, 8),
                labels = c("Couple with children", 
                           "Couple without children",
                           "Mother with children",
                           "Father with Children",
                           "Alone woman",
                           "Alone man",
                           "Several families",
                           "Other"))

```

### 2.3. Variables:

**Independent**:

hsn: Home has been rehabilitated

bld: building has been rehabilitated

env: environment has been rehabilitated

**Level**:

exp: experimental zone (vs treatment)

**Dependent variables and mechanisms**:

sths1 / sths0 / sths : satisfaction with housing before (0) and after (1) the rehabilitation and its difference (0 - 1)

stng: satisfaction with the neighbourhood

resmov: propensity for residential mobility

**Covariables**:

eco: perceived income (1-10)

family: kind of family

age: years living in the neighbourhood

apoyo: social support index (log)

relvel: participation in social activities in the neighbourhood

problemviv: housing problems index

### 2.3. Descriptive graphs

```{r, echo = FALSE, warnings = FALSE, message=FALSE}

df$hsn <- cut(df$P0209, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))
df$bld <- cut(df$P0214, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))
df$env <- cut(df$P0219, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))

df$eco <- scale(df$P1007)
df$stng <- df$P0102_A
df$sths <- df$diffviv
df$resmov <- df$P0102_C
df$support <- df$apoyo
df$sths1 <- df$P0224
df$sths0 <- df$P0225
df$eco <- df$P1007
df$age <- df$P0109
df$family <- df$P0902

means <- df %>% group_by(TIPO_ZONA) %>% summarise(mean_A = mean(P0102_C), sd_A = sd(P0102_C), n = n()) %>% ungroup()
means$se_A <- means$sd_A / sqrt(means$n)
means$lower_ci <- means$mean_A - 1.96 * means$se_A
means$upper_ci <- means$mean_A + 1.96 * means$se_A

means %>% ggplot(aes(x = TIPO_ZONA, y = mean_A)) +
  geom_bar(stat = "identity", fill = c("#F8766D", "#00BA38")) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(x = "Zones", y = "Residential Mobility Preferences (1-5)", title="Propensity of leaving the neighbourhood based on the zone") + theme_classic()

means <- df %>% 
  group_by(TIPO_ZONA) %>% 
  summarise(mean_var1 = mean(P0209), sd_var1 = sd(P0209), 
            mean_var2 = mean(P0214), sd_var2 = sd(P0214), 
            mean_var3 = mean(P0219), sd_var3 = sd(P0219), 
            n = n()) %>% 
  ungroup()

means$se_var1 <- means$sd_var1 / sqrt(means$n)
means$lower_ci_var1 <- means$mean_var1 - 1.96 * means$se_var1
means$upper_ci_var1 <- means$mean_var1 + 1.96 * means$se_var1

means$se_var2 <- means$sd_var2 / sqrt(means$n)
means$lower_ci_var2 <- means$mean_var2 - 1.96 * means$se_var2
means$upper_ci_var2 <- means$mean_var2 + 1.96 * means$se_var2


means$se_var3 <- means$sd_var3 / sqrt(means$n)
means$lower_ci_var3 <- means$mean_var3 - 1.96 * means$se_var3
means$upper_ci_var3 <- means$mean_var3 + 1.96 * means$se_var3

# Plot the means for each variable based on the group
a <- ggplot(means, aes(x = TIPO_ZONA)) +
  geom_bar(aes(y = mean_var1), stat = "identity", fill = c("#F8766D", "#00BA38"), position = "dodge") +
  geom_errorbar(aes(ymin = lower_ci_var1, ymax = upper_ci_var1), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "", y = "% of neighbors who have been rehabilitated", title="House") + ylim(0, 0.5) +
  theme_classic()
b <- ggplot(means, aes(x = TIPO_ZONA)) +
  geom_bar(aes(y = mean_var2), stat = "identity", fill = c("#F8766D", "#00BA38"), position = "dodge") +
  geom_errorbar(aes(ymin = lower_ci_var2, ymax = upper_ci_var2), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "", y = "", title="Building") + ylim(0, 0.5) +
  theme_classic()
c <- ggplot(means, aes(x = TIPO_ZONA)) +
  geom_bar(aes(y = mean_var3), stat = "identity", fill = c("#F8766D", "#00BA38"), position = "dodge") +
  geom_errorbar(aes(ymin = lower_ci_var3, ymax = upper_ci_var3), width = 0.2, position = position_dodge(0.9)) +
  labs(x = "", y = "", title="Environment") + ylim(0, 0.5) +
  theme_classic()


title <- ggdraw() + draw_label("Incidence of Rehabilitation by Zone", fontface = "bold", x = 0, hjust = -0.05)
plot_grid(title, plot_grid(a, b, c, ncol = 3, labels = "AUTO", align = "h"), legend, ncol = 1, rel_heights = c(0.1, 0.9))


means <- df %>%
  group_by(hsn, TIPO_ZONA) %>%
  summarise(
    Mean_diff = mean(diffviv),
    SE_diff = sd(diffviv) / sqrt(n())
  )

# plot means
a <- ggplot(means, aes(x = hsn, y = Mean_diff, fill = TIPO_ZONA)) +
  geom_col(position = "dodge", alpha = 0.8, color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = Mean_diff - 1.64*SE_diff, ymax = Mean_diff + 1.64*SE_diff), 
                position = position_dodge(width = 0.9), 
                width = 0.2, size = 0.5) + 
  ylab("Increase in Housing Satisfaction (0-10)") +
  xlab("House") +
  ggtitle("") + ylim(-0.05, 1.2) +
  guides(fill=guide_legend(title="Zone")) + theme_classic() +
  theme(legend.position = "top", legend.text = element_text(size = 12), legend.title = element_text(face="bold")) 


means <- df %>%
  group_by(bld, TIPO_ZONA) %>%
  summarise(
    Mean_diff = mean(diffviv),
    SE_diff = sd(diffviv) / sqrt(n())
  )

b <- ggplot(means, aes(x = bld, y = Mean_diff, fill = TIPO_ZONA)) +
  geom_col(position = "dodge", alpha = 0.8, color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = Mean_diff - 1.64*SE_diff, ymax = Mean_diff + 1.64*SE_diff), 
                position = position_dodge(width = 0.9), 
                width = 0.2, size = 0.5) + 
  ylab("") +
  xlab("Building") +
  ggtitle("") + theme_classic() + ylim(-0.05, 1.2) +
  theme(legend.position = "none")

means <- df %>%
  group_by(env, TIPO_ZONA) %>%
  summarise(
    Mean_diff = mean(diffviv),
    SE_diff = sd(diffviv) / sqrt(n())
  )

c <- ggplot(means, aes(x = env, y = Mean_diff, fill = TIPO_ZONA)) +
  geom_col(position = "dodge", alpha = 0.8, color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = Mean_diff - 1.64*SE_diff, ymax = Mean_diff + 1.64*SE_diff), 
                position = position_dodge(width = 0.9), 
                width = 0.2, size = 0.5) +
  ylab("") + 
  xlab("Environment") +
  ggtitle("") + theme_classic() + ylim(-0.05, 1.2) +
  theme(legend.position = "none")

legend <- get_legend(a)

title <- ggdraw() + draw_label("Increase in Housing Satisfaction based on rehab. and zone", fontface = "bold", x = 0, hjust = -0.05)
plot_grid(title, plot_grid(a +  theme(legend.position = "none"), b, c, ncol = 3, labels = "AUTO", align = "h"), legend, ncol = 1, rel_heights = c(0.1, 0.9))


# return variables to numerical to perform the regressions below
df$hsn <- df$P0209
df$bld <- df$P0214
df$env <- df$P0219
```



## 3. Methodology & Results

### 3.1. DiD Estimation

We have done a pseudo-DiD design in order to evaluate if the rehabilitation have an actual positive impact on housing satisfaction.

We use housing satisfaction as the mechanism through the rehabilitation affects the propensity of leaving the neighbourhood, since the rehabilitation could have not improve the households at all, and if this happens we obviously expect the effect to be 0.

We perform the DiD design with every rehabilitation variable -housing, building and environment-:


$y_{it} = \alpha + \beta_1 \cdot R_i + \beta_2 \cdot D_t + \gamma \cdot (R_i \times D_t) + \beta_3 \cdot R_i*Z_i + \beta_4 \cdot C_i + \epsilon_{it}$


where:
  $y_{it}$ represents the outcome variable for unit $i$ at time $t$ (before or after the rehabilitation).
  $R_i$ is a binary treatment indicator variable, taking a value of 1 if unit $i$ is in the rehabilited group and 0 otherwise.
  $D_t$ is a binary indicator variable for the post-treatment period, taking a value of 1 if time $t$ is in the post-rehabilitation period and 0 otherwise.
  $Z_i$ is a binary indicator variable for the areas, taking a value of 1 if time $t$ is in the affected area and 0 otherwise.
  $C_i$ represents all the confounders taking in account to avoid bias estimation of $R_i$ and $Z_i$


**Housing**
```{r, echo=FALSE}

model <- lm(sths1 ~ hsn + sths0 + hsn*sths0 + problemviv + eco + support  + exp:hsn + relvel + age + family, data = df) 
coeftest(model, vcov = vcovHC(model, type = "HC1"))
```

**Building**
```{r, echo=FALSE}

model <- lm(sths1 ~ bld + P0225 + bld*sths0 + problemviv + eco + support  + exp:bld + relvel + age + family, data = df) 
coeftest(model, vcov = vcovHC(model, type = "HC1"))
```

**Environment**
```{r, echo=FALSE}

model <- lm(sths1 ~ env + P0225 + env*sths0 + problemviv + eco + support  + exp:env + relvel + age + family, data = df) 
coeftest(model, vcov = vcovHC(model, type = "HC1"))
```

### 3.2. SEM Model

Now we evaluate the effect of rehab on the propensity for leaving the neighbourhood using a Structural Equation Modeling in order to avoid confusion between endongenous and exogenous variables and measure indirect effects. Here we do not include any covariates with non-significant effects (90% CI), to avoid confusion.


```{r}

indirind <- '
# measurement model
rehab =~ hsn + bld + env
# regressions
rehab ~ exp
sths ~ rehab + support + problemviv  
stng ~ sths + support  + eco + problemviv 
resmov ~ stng + support + eco + problemviv
'


fit <- sem(indirind, data=df)
```

**Graph of the SEM Model (only significant effects shown)**
```{r, echo=FALSE}
semPaths(fit,style = "lisrel", "std","hide") 
```

**Results of the SEM Model (only significant variables included)**
```{r, echo=FALSE}
summary(fit, standardized=TRUE, fit.measures=TRUE)
```


**Comparison of effects: Income and rehabilitation**

```{r, echo=FALSE}
lower1 = (1.569-(1.96*0.129))*(0.049-(1.96*0.008))*(-0.642+(1.96*0.014))
coef1 = 1.569*0.049*-0.642
upper1 = (1.569+(1.96*0.129))*((0.049+(1.96*0.008))*(-0.642-1.96*0.014))

# estimation of eco total effect with CI
coef2 = -0.059 + (0.076*-0.642)
lower2 <- (-0.059 + (1.96 * 0.014)) + ((0.076 + (1.96 * 0.012)) * (-0.642 + (1.96 * 0.014)))
upper2 <- (-0.059 - (1.96 * 0.014)) + ((0.076 - (1.96 * 0.012)) * (-0.642 - (1.96 * 0.014)))


coefdata <- data.frame(
  variables = c("Rehabilitation", "Income level"),
  coefficient = c(coef1, coef2),
  lower_bound = c(lower1, lower2),
  upper_bound = c(upper1, upper2),
  stringsAsFactors = FALSE
)

ggplot(coefdata, aes(y = variables, x = coefficient)) +
  geom_point(size = 4, color = "black") +
  geom_errorbarh(
    aes(xmin = lower_bound, xmax = upper_bound),
    height = 0.3,
    color = "black",
    size = 0.8
  ) +
  labs(
    x = "Coefficient and 95% CI",
    y = "Variable",
    title = "Estimation of the effects in the SEM model (standarized variables)"
  ) +
  theme_minimal() + xlim(0, -0.20)
```


### 3.3. Final Causal DAG
```{r, echo = FALSE}
mov_dag <- ggdag::dagify(mov ~ satbar,
                         satbar ~ diffviv,
                         diffviv ~ rehab,
                         rehab ~ zona,
                         labels = c("mov" = "Mobility Preferences", 
                                    "satbar" = "Neighbourhood Satisfaction",
                                    "diffviv" = "Increase of Housing Satisfaction",
                                    "rehab" = "Housing Rehabilitation",
                                    "zona" = "Experimental neighbourhoods"))

                         
ggdag::ggdag(mov_dag, # the dag object we created
             text = FALSE, # this means the original names won't be shown
             use_labels = "label") + # instead use the new names
  theme_void()
```

## 4. Summary Statistics

```{r, echo=FALSE, warnings=FALSE, error=FALSE, message=FALSE}

df$hsn <- cut(df$P0209, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))
df$bld <- cut(df$P0214, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))
df$env <- cut(df$P0219, breaks = c(-Inf, 0, Inf), labels = c("No", "Yes"))
df$eco <- df$P1007

st(df, vars = c("TIPO_ZONA", "hsn", "bld", "env", "family"))

st(df, vars = c("eco", "support", "stng", "sths", "problemviv", "resmov", "relvel", "age"))
```
